---
title: "CT-Pop Cell Type Stacked Bar Graphs Notebook"
output:
  # html_document:
  # df_print: paged
  pdf_document: 
    dev: pdf
    fig_caption: no
    number_sections: yes
    toc: yes
    toc_depth: 3
documentclass: article
urlcolor: blue
classoption: a4paper
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
---

# Environment 

```{r environment, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)

options(scipen = 999)

knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align='center',
    fig.pos='H',
    fig.path = "../output/barplots/",
    dev = c("pdf"),
    dpi=500
)
```

# Load Data

```{r load}
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v5.csv", header=T)
```

# Prepare data

```{r prep}
# Remove "sc_proteomics"
df <- df[df$tool != "sc_proteomics",]

# Prepare dataset
#df$dataset <- factor(df$dataset)

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace("right ", "") %>%
  str_replace("left ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace(" of body", "") %>%
  str_to_title() %>% 
  str_trim()
  
df$organ <- factor(df$organ)

# Prepare Cell Labels
df$cell_label <- df$cell_label %>%
  str_to_upper()

# Prepare tool
df$tool <- df$tool %>%
  str_to_title() 

df$tool <- factor(df$tool)

#ct <- unique(df$cell_label)
#write.csv(ct, file=paste0(getwd(),"/ct_pop_cellTypeList.csv"), row.names=F)
```

# Creating labels for bar plots
```{r strings}
# Extract anatomical structures from dataset URI
df$structure <- NA
df$structure[!grepl("https://doi.org/",df$dataset) & 
             grepl('#',df$dataset)] <-
  df$dataset[!grepl("https://doi.org/",df$dataset) & 
             grepl('#',df$dataset)] %>%
  str_split_i(pattern="\\$", i=-1) %>%
  str_replace_all(pattern="\\%20"," ") %>%
  str_replace(pattern="heart ","") %>%
  str_to_title()
  

# Labels for datasets
df$label <- NA

## Labels datasets without structures
df[is.na(df$structure),]$label <-
  paste0(df[is.na(df$structure),]$organ,"-",
         df[is.na(df$structure),]$dataset)

## Labels for datasets with structures 
df[!is.na(df$structure),]$label <-
  paste0(df[!is.na(df$structure),]$organ,"-",
         df[!is.na(df$structure),]$structure,"-",
         df[!is.na(df$structure),]$dataset)

# Factor label and dataset vars
df$label <- as.factor(df$label)
df$dataset <- as.factor(df$dataset)
```

# Pivot table

```{r pivot}
# Organ Cell Type Percentages
piv <- df %>% 
  select(organ, cell_label, percentage) %>%
  ddply(.(organ,cell_label), summarise,
        mean_per = mean(percentage),
        sd_per = sd(percentage))

#write.csv(piv, file=paste0(getwd(),"/ct_pop_organ_celltype_meanPercent.csv"),
#          row.names = F)

# Cell counts for each organ
piv2 <- piv %>%
  count(organ) %>%
  arrange(desc(n))
names(piv2)[2] <- "cells"
```

# Ranking dataset cell types by organ and tool

```{r rank}
df <- df %>% 
  select(label, tool, organ, cell_label, percentage) %>%
  arrange(organ, label, tool, desc(percentage)) %>% 
  group_by(label, organ, tool) %>%
  mutate(rank=row_number()) %>% 
  filter(rank <= 10)

#write.csv(df,file=paste0(getwd(),"/ct_pop_dataset_organ_top10CellTypes_tools.csv"),
#          row.names = F)

si <- df %>% 
  filter(organ=="Small Intestine") %>%
  mutate(cell_label = factor(cell_label))

li <- df %>% 
  filter(organ=="Large Intestine") %>%
  mutate(cell_label = factor(cell_label))

lng <- df %>% 
  filter(organ=="Respiratory System") %>%
  mutate(cell_label = factor(cell_label))

hrt <- df %>% 
  filter(organ=="Heart") %>%
  mutate(cell_label = factor(cell_label))

kid <- df %>% 
  filter(organ=="Kidney") %>%
  mutate(cell_label = factor(cell_label))

skin <- df %>% 
  filter(organ=="Skin") %>%
  mutate(cell_label = factor(cell_label))

ute <- df %>% 
  filter(organ=="Ureter") %>%
  mutate(cell_label = factor(cell_label))

liv <- df %>% 
  filter(organ=="Liver") %>%
  mutate(cell_label = factor(cell_label))

spl <- df %>% 
  filter(organ=="Spleen") %>%
  mutate(cell_label = factor(cell_label))

lgb <- df %>% 
  filter(organ=="Lactiferous Glands In Breast") %>%
  mutate(cell_label = factor(cell_label))

mrs <- df %>% 
  filter(organ=="Male Reproductive System") %>%
  mutate(cell_label = factor(cell_label))
```

# Stacked Bar Graphs

## Heart Datasets

```{r barplot-heart_v0.1, fig.height=40, fig.width=15, warning=FALSE}
hrt %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Heart Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Respiratory System Datasets

```{r barplot-lung_v0.1, fig.height=18, fig.width=15, warning=FALSE}
lng %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Respiratory System Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```


## Liver Datasets

```{r barplot-liver_v0.1, fig.height=4, fig.width=15, warning=FALSE}
liv %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Liver Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Lactiferous Glands In Breast Datasets

```{r barplot-lgb_v0.1, fig.height=4, fig.width=15, warning=FALSE}
lgb %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Lactiferous Glands In Breast Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Kidney Datasets

```{r barplot-kidney_v0.1, fig.height=20, fig.width=15, warning=FALSE}
kid %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Kidney Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Large Intestine Datasets

```{r barplot-largeIntestine_v0.1, fig.height=11, fig.width=15, warning=FALSE}
li %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Large Intestine Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Small Intestine Datasets

```{r barplot-smallIntestine_v0.1, fig.height=11, fig.width=15, warning=FALSE}
si %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Small Intestine Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Skin Datasets

```{r barplot-skin_v0.1, fig.height=4, fig.width=15, warning=FALSE}
skin %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Skin Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Spleen Datasets

```{r barplot-spleen_v0.1, fig.height=5, fig.width=15, warning=FALSE}
spl %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Spleen Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Ureter Datasets

```{r barplot-ureter_v0.1, fig.height=5, fig.width=15, warning=FALSE}
ute %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Ureter Datasets - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```

## Male Reproductive System Datasets

```{r barplot-mrs_v0.1, fig.height=4, fig.width=15, warning=FALSE}
mrs %>% 
  ggplot(aes(y=label, x=percentage)) +
  geom_col(aes(fill=cell_label)) +
  facet_grid(rows = vars(tool)) +
  scale_y_discrete(expand=c(0,0)) + 
  scale_x_continuous(expand=c(0,0)) + 
  guides(fill="none") +
  labs(title="Male Reproductive System - Top 10 Cell Types, by Percentage & Tool",
       x="Percent",y="Dataset")
```