---
title: "Organ, Anatomical Structures & Cell Type Bi-modal Networks - Notebook"
output:
  html_document
---

# Environment 

```{r environment, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(igraph)
library(ggraph)
library(ggnetwork)
library(stringr)

options(scipen = 999)
```

# Load Data

This section of script identifies the unique identifiers, organ, and anatomical structure associated with a data set. The identifiers are used to generate a url that is used to download the cell type measurements for each data set. The downloaded data are combined into a single data frame for further data preparation.

```{r load}
# Data - bar graph 
# Includes organ, anatomical structure, cell types measures by tool and donor gender
data_set_meta <- read.csv(paste0("../network/HRApop_intestine_datasets-ASregion_crosswalk-Hickey.csv"), 
                          header=T, fileEncoding = "UTF8")

# Data set id for cell type measures
data_set_ids <- unique(data_set_meta$dataset_id)
#data_hm <- data
# data frame for downloaded results
data <- data.frame(id = as.character(),
                   #organ = as.character(),
                   #AS = as.character(),
                   cell_type = as.character(),
                   count = as.numeric())

# URL for collecting data set measurements
data_url <- "https://raw.githubusercontent.com/cns-iu/hra-ct-summaries-mx-spatial-data/main/intestine_omap_2_0001/cell_type_counts/"

# For loop to collect data with cell type measures
for(i in 1:length(data_set_ids)){
  dataset_id <- data_set_ids[i]
  tmp <- read.csv(paste0(data_url,data_set_ids[i],".csv"), header = T)
  tmp <- cbind(dataset_id,tmp)
  data <- rbind(data,tmp)
  rm(dataset_id,tmp)
}

# Join CT measurements with data set metadata
data <- left_join(data,data_set_meta, by="dataset_id") %>%
           select(dataset_id, organ, uberon_id, as_uberon_label, 
                  cell_type, count) 
names(data) <- c("id","organ","as_id","as","cell_label","count")

# Factoring data
data$id <- factor(data$id)
data$as <- factor(str_to_title(data$as))

# Clean-up 
rm(i, data_set_ids, data_url)
```

# Prepare data

## Calculate cell type counts

This section of the script is used to calculate the number of cells identified
in data sets associated with an organ, their anatomical structures, and cell 
types. First, we calculate the total count of cells in a data set, which are 
associated with organs and anatomical structures by the team collecting a sample.
Next, we calculate number of cells for a given cell type found in an organ and 
anatomical structure, as well as the mean cell counts, and the number of data set
found in the measurements. Finally, we combine the two cell count calculations 
into a single data frame, and then calculate the percentage each cell type 
contributes to data collected from each anatomical structure.

```{r cell_type_pivots}
# Aggregate data to get limited set of nodes and edges.
# Organ-AS Total Cell Count (overall)
data_total_cell_count <- data %>% 
                          ddply(.(organ, as_id, as), summarise,
                                  total_cell_count = sum(count))
# Organ-AS-CT counts 
data_agg <- data %>% 
              ddply(.(organ, as_id, as, cell_label), summarise,
                    datasets = length(unique(id)),
                    cell_count = sum(count),
                    mean_cell_count = median(count))

# Combine aggregated data - organ - as - ct counts and percentages.
data <- left_join(data_agg,data_total_cell_count, by=c("organ","as_id","as")) %>%
          select(organ,as_id,as,cell_label,datasets,mean_cell_count,cell_count,total_cell_count) %>%
          mutate(percentage = round(cell_count/total_cell_count,5))

rm(data_agg,data_total_cell_count)

data$tool <- "Spatial"

# Export results
write.csv(data,"../heatmaps/HRApop_Spatial_OASCT_CellCount_Percentage.csv",
          row.names = F)
write.csv(data,"../network/HRApop_Spatial_OASCT_CellCount_Percentage.csv",
          row.names = F)
```

## Network Extraction

This section of code is used to create networks from the validation data set. First data is subset into sets based on the cell type annotation tool used for an observation, and added to a list object that will be used in a *for loop*.

The for loop processes each tool subset to create a node list of organs, anatomical structures and cell types and an edge list that link organs to anatomical structures, and anatomical structures to cell types, creating a tree-like multi-modal network graph. After creating a node and edge list for a given data subset, results are saved as CSV files and exported to the R global environment as igraph networks from data.frames.

```{r nets_extract}
## Subset data for graphs, based on CTann tool
df_list <- list(data[data$tool=="Spatial",])

## Loops through df_list to create network graphs, for each CTann tool.
for(i in 1:length(df_list)){
    ## Node list creation
    anatomical_structures <-
      df_list[[i]] %>% 
      select(organ, as, datasets, total_cell_count) %>%
      mutate(id_lab = as,
             type = "Anatomical Structure") %>% 
      dplyr::rename(label = as,
                    partOf = organ,
                    records = datasets,
                    cell_count = total_cell_count) %>%
      select(id_lab, label, type, partOf, 
             records, cell_count) %>%
      distinct()
    
    cell_types <- 
      df_list[[i]] %>% 
      select(organ, cell_label, datasets, cell_count) %>%
      ddply(.(cell_label), summarise,
            label = head(cell_label)[1],
            records = sum(datasets),
            cell_count = sum(cell_count),
            partOf = length(unique(organ))) %>%
      mutate(id_lab = cell_label,
             type = "Cell Type",
             partOf = paste0("Organ(s): ",partOf)) %>% 
      select(id_lab, label, type, partOf, 
             records, cell_count)
    
    # Combine node lists into one list
    nodes <- rbind(anatomical_structures, cell_types)
    rm(anatomical_structures, cell_types)
    nodes <- 
      nodes %>% 
      mutate(id = as.numeric(row.names(nodes))-1) %>%
      select(id, id_lab, label, type, partOf, 
             records, cell_count)
    
    # Create Edge List
    edges <- 
      df_list[[i]] %>% 
      ddply(.(as, cell_label), summarise,
            weight_recs = sum(datasets),
            percentage = mean(percentage)) %>%
      mutate(source_id_lab = as,
             target_id_lab = cell_label,
             weight = 1) %>% 
      select(source_id_lab, target_id_lab, 
             weight, weight_recs, percentage) %>%
      distinct()
    
    # Update edge list source and target ids
    edges <- left_join(edges, 
                       nodes[,1:2], by=c("source_id_lab"="id_lab")) %>%
             dplyr::rename(source = id)
    edges <- left_join(edges, 
                       nodes[,1:2], by=c("target_id_lab"="id_lab")) %>%
             dplyr::rename(target = id) %>%
             select(source, target, source_id_lab, target_id_lab,
                    weight, weight_recs, percentage)
    
    ## Create network from data_frames
    ## Save edge and node lists as data files
    write.csv(edges, file=paste0("../network/HRApop_OASCT_spatial_edgeList.csv"),
              row.names = FALSE)
    write.csv(nodes, file=paste0("../network/HRApop_OASCT_spatial_nodeList.csv"),
              row.names = FALSE)

    # Re-nameing for igraph
    edges <- edges %>%
              dplyr::rename(from = source,
                            to = target)    
    
    # Combines network edge and node lists
    network <- igraph::graph_from_data_frame(edges, directed=T,
                                  vertices=nodes)
    # Reassigns name of network object with tool names 
    rm(nodes,edges)
}

rm(df_list, i)
```
