---
title: "Organ, Anatomical Structures & Cell Type Bi-modal Networks - Notebook"
output:
  html_document
---

# Environment 

```{r environment, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(igraph)
library(ggraph)
library(ggnetwork)
library(stringr)

options(scipen = 999)
```

# Load Data

```{r load}
# Data - bar graph 
# Includes organ, anatomical structure, cell types measures by tool and donor gender
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v6.csv", 
                 header=T, fileEncoding = "UTF8")

# Cell Type Crosswalks - Data Cleaning
# https://github.com/hubmapconsortium/hra-workflows-runner/tree/main/crosswalking-tables
# New version June 2024
## Azimuth
azimuth <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/azimuth/v1.0/assets/azimuth-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## Cell Typist
celltypist <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/celltypist/v1.0/assets/celltypist-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## PopV
popv <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/popv/v1.0/assets/popv-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)
```

# Prepare data

```{r prep}
# Prepare crosswalk field names
names(celltypist) <- names(popv) <- names(azimuth) <- c("organ_level","organ_id",
                                                        "annotation_label","annotation_label_id",
                                                        "cell_label","cell_id","cell_match")

# Title Case for Annotation Labels
popv$annotation_label <- str_to_title(popv$annotation_label)
azimuth$annotation_label <- str_to_title(azimuth$annotation_label)
celltypist$annotation_label <- str_to_title(celltypist$annotation_label)

# Add record number to data 
df$rec <- as.numeric(row.names(df))

# Prepare Sex
df$sex <- factor(df$sex)

# Prepare Tool
df$tool <- factor(df$tool, 
                  levels=c("azimuth","celltypist","popv","sc_proteomics"),
                  labels=c("Azimuth","CellTypist","popV","SC Proteomics"))

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace("left ", "") %>%
  str_replace("right ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace("left ", "") %>%
  str_replace(" of body", "") %>% 
  str_replace("respiratory system","lung") %>%
  str_replace("male reproductive system","prostate") %>%
  str_to_title()
df$organ <- factor(df$organ)

# Prepare anatomical structure labels
df$as <- df$as %>% 
  str_to_title() %>%
  str_replace(" Part "," part ") %>%
  str_replace(" Of "," of ") %>% 
  str_replace(" In "," in ") %>%
  str_replace("Heart ","")

df$cell_label <- df$cell_label %>%
  str_replace("_"," ")

# Common labels for data frames.
labels <-  c("sex", "tool", "modality", "organ", "organId", 
             "as", "as_id",  "cell_id", "cell_label", "cell_count",
             "percentage", "asct_relation_in_asctb_table",
             "indirect_asct_relation_in_asctb_table", 
             "annotation_label", "annotation_label_id", "cell_label2")
```

## Clean up Cell Labeling

```{r cell_type_labels}
## Heart - 1086
heart <- df[df$organ==levels(df$organ)[1],]
# Updating cell labels using crosswalks
tmp1 <- 
  join(heart[heart$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Heart_L2" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
tmp2 <- 
  join(heart[heart$tool=="popV",c(1:11)],
       popv[popv$organ_level=="heart" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left") 
tmp3 <- 
  join(heart[heart$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="Healthy_Adult_Heart_pkl" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Clean up names
names(tmp1) <- names(tmp2) <- names(tmp3) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Combine temp files
heart <- rbind(tmp1,tmp2,tmp3) 
# Add back missing cell labels when annotation label is missing
heart[is.na(heart$annotation_label),]$cell_label2 <- 
  heart[is.na(heart$annotation_label),]$cell_label


## Lungs - 826
lung <- df[df$organ==levels(df$organ)[6],]
# Updating cell labels using crosswalks
tmp1 <- 
  join(lung[lung$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Lung_v2_finest_level" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
tmp2 <- 
  join(lung[lung$tool=="popV",c(1:11)],
       popv[popv$organ_level=="lung" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
tmp3 <- 
  join(lung[lung$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="Human_Lung_Atlas_pkl" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
#Clean up names
names(tmp1) <- names(tmp2) <- names(tmp3) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Combine temp files
lung <- rbind(tmp1,tmp2,tmp3) 
# Add back missing cell labels when annotation label is missing
lung[is.na(lung$annotation_label),]$cell_label2 <- 
  lung[is.na(lung$annotation_label),]$cell_label
# write.csv(lung,
#           file=paste0("../bargraphs/HRApop_Lung_AS-CTBarGraph-cellID_annotationIDs.csv"),
#           row.names = FALSE)

## Kidney - 920
kidney <- df[df$organ==levels(df$organ)[2],]
# Updating cell labels using crosswalks
kidney <- 
  join(kidney[kidney$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Kidney_L3" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
#Clean up names
names(kidney) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Add back missing cell labels when annotation label is missing
kidney[is.na(kidney$annotation_label),]$cell_label2 <- 
  kidney[is.na(kidney$annotation_label),]$cell_label
# write.csv(kidney,
#           file=paste0("../bargraphs/HRApop_Kidney_AS-CTBarGraph-cellID_annotationIDs.csv"),
#           row.names = FALSE)


## Large Intestine - 1790
largeIntestine <- df[df$organ==levels(df$organ)[4],]
# Updating cell labels using crosswalks
tmp1 <- 
  largeIntestine[largeIntestine$tool=="SC Proteomics",c(1:11)] %>%
  mutate(annotation_label=NA,
         annotation_label_ids=NA,
         cell_label2=NA)
tmp2 <- 
  join(largeIntestine[largeIntestine$tool=="popV",c(1:11)],
       popv[popv$organ_level=="large intestine" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
tmp3 <- 
  join(largeIntestine[largeIntestine$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="intestine_L1" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
#Clean up names
names(tmp1) <- names(tmp2) <- names(tmp3) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Combine temp files
largeIntestine <- rbind(tmp1,tmp2,tmp3) 
# Add back missing cell labels when annotation label is missing
largeIntestine[is.na(largeIntestine$annotation_label),]$cell_label2 <- 
  largeIntestine[is.na(largeIntestine$annotation_label),]$cell_label


## Small Intestine
smallIntestine <- df[df$organ==levels(df$organ)[9],]
# Updating cell labels using crosswalks
tmp1 <- 
  smallIntestine[smallIntestine$tool=="SC Proteomics",c(1:11)] %>%
  mutate(annotation_label=NA,
         annotation_label_ids=NA,
         cell_label2=NA)
tmp2 <- 
  join(smallIntestine[smallIntestine$tool=="popV",c(1:11)],
       popv[popv$organ_level=="small intestine" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left") 
tmp3 <- 
  join(smallIntestine[smallIntestine$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="intestine_L1" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
#Clean up names
names(tmp1) <- names(tmp2) <- names(tmp3) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Combine temp files
smallIntestine <- rbind(tmp1,tmp2,tmp3) 
# Add back missing cell labels when annotation label is missing
smallIntestine[is.na(smallIntestine$annotation_label),]$cell_label2 <- 
  smallIntestine[is.na(smallIntestine$annotation_label),]$cell_label


## Combine data frames into a single data frame
df2 <- 
  rbind(heart,lung,kidney,largeIntestine,smallIntestine) %>%
  arrange(organ,tool,sex,percentage)
df2$annotation_label <- gsub("[φ]", "ph", df2$annotation_label)

# Add in missing cell labels for cell_label2 
df2[is.na(df2$cell_label2),]$cell_label2 <- 
  df2[is.na(df2$cell_label2),]$cell_label
 
## Export annotation look-up
organ_as_ct_lookup <- 
  rbind(heart,lung,kidney,largeIntestine,smallIntestine)
# write.csv(organ_as_ct_lookup, 
#           file=paste0("../bargraphs/HRApop_AS-CTBarGraph-cellID_annotationIDs.csv"),
#           row.names = FALSE)

## Re-aggregate data to remove duplication caused by sex.
df2 <- df2 %>%
  ddply(.(tool, organ, organId, as, as_id, cell_id, cell_label2), summarise,
        cell_count = mean(cell_count),
        percentage = mean(percentage))

# Clean up
rm(tmp1,tmp2,tmp3)
rm(heart,lung,kidney,largeIntestine,smallIntestine)
```

### Identification of Representative Cell Types Using Term Frequency - Inverse Document Frequency  (TF-IDF) Analysis

The goal of this analysis is to identify cell types that are most descriptive of
an tool-organ-anatomical structure set. 

This section implements a term-frequency & inverse document frequency analysis to the
data set to identify cell types most representative of the combinations of organs, & anatomical structures. In this analysis the sets of tools, organs, & anatomical structures are equivalent to documents, and cell types are equivalent to terms in a document. 

In a TF-IDF analysis TF is the frequency of a term divided by the total number of terms
in a document. In this case, cell type frequency (cTF) would be equivalent to the cell type *percentage* measurement, which is the number of cells (cell count) divided by the total cells found in a given tool-organ-as tissue sample.

```{r tfidf}
# Initial data selection - cell type term frequencies
tf <- 
  df2 %>% 
  # removes duplicate data rows
  distinct() %>%
  # combines sex and tools as a record sum of cell count and avg percentage
  ddply(.(organ, as, cell_id, cell_label2), summarise,
        records = length(cell_count),
        cell_count = sum(cell_count),
        tf = mean(percentage))

tf[is.na(tf$cell_label2),]$cell_label2 <- tf[is.na(tf$cell_label2),]$cell_label

# Calculate IDF based on TF data.
## Number of documents (unique combinations of tools, organ, anatomical structures)
docs <- tf %>% 
          select(organ, as) %>%
          distinct() %>%
          nrow()

## Calculate IDF 
idf <- tf %>% 
          group_by(organ, as) %>%
          ddply(.(cell_id, cell_label2), summarise,
                term_docs = length(records)) %>%
          mutate(idf = log(docs/term_docs)) %>%
          arrange(desc(idf))

# Join IDF results to the TF data
tf_idf <- 
  left_join(tf, idf,by=c("cell_id","cell_label2")) %>%
  mutate(tfidf = tf*idf) %>% 
  select(organ, as, cell_id, cell_label2, 
         cell_count, tf, idf, tfidf) %>%
  arrange(organ,as,desc(tfidf))

# Add TF IDF results to data for bar graphs.
df2 <- 
  left_join(df2, tf_idf[,c(1:4,6:8)],
            by=c("organ","as","cell_id","cell_label2"))

# Rename cell_label2 to cell_label
names(df2)[7] <- "cell_label"

rm(tf,idf,docs)
```

## Network Extraction

This section of code is used to create networks from the validation data set. First data is subset into sets based on the cell type annotation tool used for an observation, and added to a list object that will be used in a *for loop*.

The for loop processes each tool subset to create a node list of organs, anatomical structures and cell types and an edge list that link organs to anatomical structures, and anatomical structures to cell types, creating a tree-like multi-modal network graph. After creating a node and edge list for a given data subset, results are saved as CSV files and exported to the R global environment as igraph networks from data.frames.

```{r nets_extract}
## Subset data for graphs, based on CTann tool
df_list <- list(df2[df2$tool=="Azimuth",],
                df2[df2$tool=="popV",],
                df2[df2$tool=="CellTypist",],
                df2[df2$tool=="SC Proteomics",])

## Loops through df_list to create network graphs, for each CTann tool.
for(i in 1:length(df_list)){
    ## Node list creation
    anatomical_structures <-
      df_list[[i]] %>% 
      select(tool, organ, as_id, as, cell_count) %>%
      ddply(.(tool, organ, as_id, as), summarise,
            records = length(cell_count),
            cell_count = sum(cell_count)) %>%
      mutate(id_lab = paste0(as_id," ",as),
             type = "Anatomical Structure") %>% 
      dplyr::rename(label = as,
                    partOf = organ) %>%
      select(tool, id_lab, label, type, partOf, 
             records, cell_count)
    
    cell_types <- 
      df_list[[i]] %>% 
      select(tool, organ, cell_id, cell_label, cell_count) %>%
      ddply(.(tool, cell_id), summarise,
            label = head(cell_label)[1],
            records = length(cell_count),
            cell_count = sum(cell_count),
            partOf = length(unique(organ)))  %>%
      mutate(id_lab = cell_id,
             type = "Cell type",
             partOf = paste0("Organ(s): ",partOf)) %>% 
      select(tool, id_lab, label, type, partOf, 
             records, cell_count)
    
    # Combine node lists into one list
    #nodes <- rbind(organs, anatomical_structures, cell_types)
    #rm(organs, anatomical_structures, cell_types)
    nodes <- rbind(anatomical_structures, cell_types)
    rm(anatomical_structures, cell_types)
    nodes <- 
      nodes %>% 
      mutate(id = as.numeric(row.names(nodes))-1) %>%
      select(id, id_lab, label, tool, type, partOf, 
             records, cell_count)

    #as_celltypes <- 
    edges <- 
      df_list[[i]] %>% 
      ddply(.(tool, as_id, as, cell_id), summarise,
            weight_recs = length(tool),
            weight_tfidf = median(tfidf)) %>%
      mutate(source_id_lab = paste0(as_id," ",as),
             target_id_lab = cell_id,
             weight = 1) %>% 
      select(source_id_lab, target_id_lab, 
             weight, weight_recs, weight_tfidf) %>%
      distinct()
    
    # Update edge list source and target ids
    edges <- left_join(edges, 
                       nodes[,1:2], by=c("source_id_lab"="id_lab")) %>%
             dplyr::rename(to = id)
    edges <- left_join(edges, 
                       nodes[,1:2], by=c("target_id_lab"="id_lab")) %>%
             dplyr::rename(from = id) %>%
             select(to, from, source_id_lab, target_id_lab,
                    weight, weight_recs, weight_tfidf)
    
    ## Create network from data_frames
    # Select tool name for graph
        tool <- 
      as.character(unique(df_list[[i]]$tool))
    
    ## Save edge and node lists as data files
    write.csv(edges, file=paste0("../network/HRApop_OASCT_",
                                 str_replace(str_to_lower(tool)," ","_"),
                                 "_edgeList.csv"),
              row.names = FALSE)
    write.csv(nodes, file=paste0("../network/HRApop_OASCT_",
                                 str_replace(str_to_lower(tool)," ","_"),
                                 "_nodeList.csv"),
              row.names = FALSE)

    # Combines network edge and node lists
    network <- igraph::graph_from_data_frame(edges, directed=T,
                                  vertices=nodes)
    # Reassigns name of network object with tool names 
    assign(paste0("network_",
                  str_replace(str_to_lower(tool[1])," ","_")),
           network, envir = .GlobalEnv)
    rm(nodes,edges,network, tool)
}

rm(df_list, i)
```
