---
title: "Kidney Anatomical Structures - CT-Pop Cell Types Stacked Bar Graphs Example"
output: html_notebook
---

### Environment 

```{r environment, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)

options(scipen = 999)

knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align='center',
    fig.pos='H',
    fig.path = "../output/barplots/",
    dev = c("pdf"),
    dpi=500
)
```

### Load Data

```{r load, echo=TRUE}
# Data - bar graph 
# Includes organ, anatomical structure, cell types measures by tool and donor gender
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v6.csv", 
                 header=T, fileEncoding = "UTF8")

# Cell Type Crosswalks - Data Cleaning
# https://github.com/hubmapconsortium/hra-workflows-runner/tree/main/crosswalking-tables
# New version June 2024
## Azimuth
azi <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/azimuth.csv",
                header=T, fileEncoding = "UTF8", skip=10)
```

### Preparing Data

```{r prep, echo=TRUE}
# Prepare crosswalk field names
names(azi) <- c("organ_level","organ_id",
                "annotation_label","annotation_label_id",
                "cell_label","cell_id","cell_match")

# Title Case for Annotation Labels
azi$annotation_label <- str_to_title(azi$annotation_label)

# Add record number to data 
df$rec <- as.numeric(row.names(df))

# Prepare Sex
df$sex <- factor(df$sex)

# Prepare Tool
df$tool <- factor(df$tool, 
                  levels=c("azimuth","celltypist","popv","sc_proteomics"),
                  labels=c("Azimuth","CellTypist","popV","SC Proteomics"))

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace("left ", "") %>%
  str_replace("right ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace("left ", "") %>%
  str_replace(" of body", "") %>% 
  str_replace("respiratory system","lung") %>%
  str_replace("male reproductive system","prostate") %>%
  str_to_title()
df$organ <- factor(df$organ)

# Prepare anatomical structure labels
df$as <- df$as %>% 
  str_to_title() %>%
  str_replace(" Part "," part ") %>%
  str_replace(" Of "," of ") %>% 
  str_replace(" In "," in ") %>%
  str_replace("Heart ","")

df$cell_label <- df$cell_label %>%
  str_replace("_"," ")
```

### Visualization

Base design for 100% stacked bar graphs representing the cell types commonly identified in anatomical structures within donated organs.

Bar graph definition:

Stacked bar represent top cell types associated with an organ anatomical structure
* Data threshold: 8% minimum cell percentage values (or mean cell percentage values PER AS)
* Bars - anatomical structure categories.
  * Bar segment-area-Cell Type percentage measures
    * Bar/Bar segments should total to at most 1 or 100%.
    * Mean cell type percentage calculation-tool, sex organ, as
  * Bar segment-color-cell types categories.
* Facets represent combination of analytic tool and gender.

#### Kidney - Anatomical Structure - Cell Types

```{r bar_kidney_as_1, fig.height=6, fig.width=14}
# Select initial data
kid <- df[df$organ==levels(df$organ)[2],] %>%  #kidney
       #df[df$organ==levels(df$organ)[6],] %>%  #lung
       #df[df$organ==levels(df$organ)[1],] %>%  #heart
  filter(sex == "Male") %>%
  filter(asct_relation_in_asctb_table == "true" | 
         indirect_asct_relation_in_asctb_table == "true") %>%
  arrange(as, desc(percentage)) %>%
  distinct() %>%
  group_by(as) %>%
  mutate(rec = row_number()) %>%
  ungroup() %>%
  arrange(cell_id)
 
#Prepare cell type labels - Kidney
kid <- join(kid, azi[azi$organ_level=="Kidney_L2" &
                 azi$cell_match=="skos:exactMatch" &
                 azi$annotation_label_id != "AZ:0000033", c(3:6)],
             by="cell_id", type="left")

# # Prepare cell type labels - Lung
# tmp1 <- join(kid[kid$tool=="Azimuth",],
#              azi[azi$organ_level=="Lung_v2_L5" &
#                  azi$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")
# tmp2 <- join(kid[kid$tool=="popV",],
#              pop[pop$organ_level=="lung" &
#                  pop$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")
# tmp3 <- join(kid[kid$tool=="CellTypist",],
#              ctp[ctp$organ_level=="Human_Lung_Atlas_pkl" &
#                  ctp$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")

# # Prepare cell type labels - heart
# tmp1 <- join(kid[kid$tool=="Azimuth",],
#              azi[azi$organ_level=="Heart_L2" &
#                  azi$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")
# tmp2 <- join(kid[kid$tool=="popV",],
#              pop[pop$organ_level=="heart" &
#                  pop$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")
# tmp3 <- join(kid[kid$tool=="CellTypist",],
#              ctp[ctp$organ_level=="heart_L1" &
#                  ctp$cell_match=="skos:exactMatch", c(3:6)],
#              by="cell_id", type="left")
# tmp3[is.na(tmp3$annotation_label),]$annotation_label <-
#   tmp3[is.na(tmp3$annotation_label),]$cell_label

# Combine temp files
#kid <- rbind(tmp1,tmp2,tmp3) 

# Add in missing values for annotation labels
names(kid)[17] <- "cell_label2"
kid[is.na(kid$annotation_label),]$annotation_label <-
  kid[is.na(kid$annotation_label),]$cell_label

# Aggregate data
kid <- kid %>%
  select(organ,as,cell_id,annotation_label,
         cell_count,percentage,tool,sex,
         asct_relation_in_asctb_table,
         indirect_asct_relation_in_asctb_table) %>%
  mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  ddply(.(asct_relation_in_asctb_table,
         indirect_asct_relation_in_asctb_table,
         organ,sex,tool,as,cell_lab), summarise,
        ct = length(cell_count),
        med_cell_ct = median(cell_count),
        med_cell_pt = median(percentage)) %>%
  arrange(sex, tool, as, desc(med_cell_pt)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) # %>%
  #filter(mean_cell_pt >= .08)

# Reorganize columns
kid <- kid %>% select(tool,sex,organ,as,cell_lab,
                      ct,med_cell_ct,med_cell_pt,
                      asct_relation_in_asctb_table,
                      indirect_asct_relation_in_asctb_table)

write.csv(kid,"HRApop-m-kidney-as-ct-medianPercentage.csv", row.names = FALSE)

# Factor Kidney AS
kid$as <- levels(factor(kid$as))

# Factor Kidney Cell Types
#View(as.data.frame(levels(factor(kid$cell_lab))))
kid$cell_lab <- factor(kid$cell_lab, 
                       levels=c("CL:1000714: Cortical Collecting Duct Principal",
                                "CL:1000716: Outer Medullary Collecting Duct Principal",
                                "CL:1000718: Inner Medullary Collecting Duct",
                                "CL:1001108: Medullary Thick Ascending Limb",
                                "CL:1001109: Cortical Thick Ascending Limb",
                                "CL:1001107: Ascending Thin Limb",
                                "CL:4030012: Descending Thin Limb Type 1",
                                "CL:4030014: Descending Thin Limb Type 3",
                                "CL:1000768: Connecting Tubule",
                                "CL:4030016: Distal Convoluted Tubule Type 1",
                                "CL:4030009: Proximal Tubule Epithelial Segment 1",
                                "CL:4030022: Medullary Fibroblast"))



color_pal_kid <- c("#CD8490","#A26771","#C0BCC8","#AADCDF","#88B1B3","#556E6F",
                   "#8DC599","#6F9A78","#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC","#675868","#E97B74","#E07770","#BE655F",
                   "#7A413D","#582F2C","#A1ACD2","#E9E4F3","#CD8490")

color_pal_kid2 <- c("#7495AF","#70A5A8","#AADCDF","#D6B6D7","#A295C9",
                    "#9C534E","#E97B74","#EDB8AC","#8DC599","#6F9A78","#F9CE8D","#A1ACD2",
                    "#556E6F","#FF0043","#BAD2E7","#D8D7DB","#4B4B5E","#C4C5C6",
                    "#A2A3A4","#818182","#5F6060","#3D3E3E","#cdcecf")

kid_bg <- kid %>% arrange(sex,tool,as,mean_cell_pt) %>%
              ggplot(aes(x=as,y=mean_cell_pt)) +
                geom_bar(aes(fill=cell_lab), stat="identity") +
                facet_wrap(facets=vars(tool, sex)) +
                labs(x="Anatomical Structure", y="Percentage (%)",
                     title="Kidney Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
                scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
                scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
                theme_light() +
                theme(legend.position = "bottom",
                      legend.title = element_text(face="bold", size=10),
                     # legend.key.spacing.y = unit(.1)
                      strip.background = element_rect(fill = "#cdcecf"),
                      strip.text = element_text(color="black", size=10, hjust=0))

# Export pdf
ggsave(file=paste0("../output/barplots/kidney_as_celltype_bargraph_0.2.pdf"),
       kid_bg, width=14, height=6, units="in")

kid_bg
```